---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300
)
```
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/poissonconsulting/fwabc.svg?branch=master)](https://travis-ci.org/poissonconsulting/fwabc)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/fwabc?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/fwabc)
[![Coverage status](https://codecov.io/gh/poissonconsulting/fwabc/branch/master/graph/badge.svg)](https://codecov.io/github/poissonconsulting/fwabc?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)

# fwabc

`fwabc` is an R package to read data from the [BC Freshwater Atlas](https://www2.gov.bc.ca/assets/gov/data/geographic/topography/fwa/fwa_user_guide.pdf). 

## Installation

Install the latest development version of `bcdata` from [GitHub](https://github.com/bcgov/bcdata) with `remotes` package:
```
install.packages("remotes")
remotes::install_github("bcgov/bcdata")
```

Install the latest development version of `fwabc` from [GitHub](https://github.com/poissonconsulting/fwabc)
```
remotes::install_github("poissonconsulting/fwabc")
```

## Usage
### Function families
+ `fwa_read_` - read features from BC Freshwater Atlas layers.
+ `fwa_pull_` - return a modified version of some input (e.g. another input type, or tributaries of that input).
+ `fwa_search_` - match a regular expression to official named features (`GNIS_NAME` and `WATERSHED_GROUP_NAME`).

### Read 
Read features from available layers using the `fwa_read()` function:
```{r stream}
library(ggplot2)
library(magrittr)
library(fwabc)

fwa_read(c("SKGT", 356439092), layer = "stream-network") 

```
All `fwa_read_` functions return a [sf](https://github.com/r-spatial/sf) object and require some combination of `WATERSHED_KEY` and `WATERSHED_GROUP_CODE` as input.

```{r lookup}
fwa_lookup_layer
```

Read features from multiple layers for Skagit River watershed group:
```{r layers}
layers <- c("stream-network", "rivers", "lakes", "watershed-groups")
x <- lapply(layers, function(x) fwa_read("SKGT", layer = x))
names(x) <- layers

ggplot() + 
    geom_sf(data = x[["watershed-groups"]], size = 0.2) +
    geom_sf(data = x[["lakes"]], size = 0.3, fill = "steelblue") +
    geom_sf(data = x[["rivers"]], size = 0.3, fill = "steelblue") +
    geom_sf(data = x[["stream-network"]][x[["stream-network"]]$STREAM_ORDER > 2,], size = 0.07) 
  
```

There is a convenience function for each layer, e.g. `fwa_read_stream_network()`, `fwa_read_watershed_groups()`, `fwa_read_coastlines()` etc.
To read an entire layer, leave the default `x = NULL`. Some layers (e.g. `stream-network` and `watersheds`) are very large.

```{r wsgroup}
library(rmapshaper)
x <- fwa_read_watershed_groups(ask = FALSE) %>%
  # simplify with rmapshaper package
  rmapshaper::ms_simplify()  

ggplot() + 
  geom_sf(data = x, size = 0.05, aes(fill = WATERSHED_GROUP_NAME), show.legend = FALSE) 
```

### Search and pull
`fwa_search_` and `fwa_pull_` functions can be used to find `WATERSHED_KEY` or `WATERSHED_GROUP_CODE`.

Match a regular expression to an official stream name (`GNIS_NAME`) or watershed group name (`WATERSHED_GROUP_NAME`):

```{r search}
fwa_search_gnis("skagi|scagi")
fwa_search_watershed_group("skagi|scagi")
```

Pull the `WATERSHED_KEY` or `WATERSHED_GROUP_CODE` from the official name and provide to `fwa_read_`:

```{r pipe}
stream <- fwa_search_gnis("skagi|scagi") %>%
  fwa_pull_watershed_key() %>%
  fwa_read_stream_network()

wshed <- fwa_search_watershed_group("skagi|scagi") %>%
  fwa_pull_watershed_group_code() %>%
  fwa_read_watershed_groups()

ggplot() +
  geom_sf(data = wshed, size = 0.2) +
  geom_sf(data = stream) 
  
```

### Tributaries
Use `fwa_pull_tributaries()` to get tributaries from a `WATERSHED_KEY`:

```{r tribs}
tribs <- fwa_search_gnis("skagit") %>%
  fwa_pull_watershed_key() %>%
  fwa_pull_tributaries(order = 1L) %>%
  fwa_read_stream_network()

ggplot() +
  geom_sf(data = wshed, size = 0.2) +
  geom_sf(data = stream) +
  geom_sf(data = tribs, size = 0.1) 

```

### Lookup tables
`fwabc` provides three lookup tables (`fwa_lookup_gnis`, `fwa_lookup_watershed_group`, `fwa_lookup_layer`), which can be useful for finding layer names, named features or which layers have data for particular features. 

## Other packages
`fwabc` read functions use the [`bcdata`](https://github.com/bcgov/bcdata) package (specifically, `bcdata::bcdc_query_geodata()`) to query the freshwater atlas remotely. `fwabc` provides convenience for common requests

```{r bcdata, eval = FALSE}
fwa_read_stream_network(c("SKGT", 356439092))

#### ---- returns the same result as ---- ####

bcdata::bcdc_query_geodata("freshwater-atlas-stream-network") %>%
  bcdata::filter(WATERSHED_GROUP_CODE == "SKGT" | WATERSHED_KEY == 356439092) %>%
  bcdata::collect()

```

and resources/functions to get tributaries and official `GNIS_NAME`, `WATERSHED_KEY` and `WATERSHED_GROUP_CODE`. 

However, if you are very familiar with the atlas and require greater filtering flexibility (or even want to use your own CQL), then we suggest trying the fantastic `bcdata` package.

## Contribution

Please report any [issues](https://github.com/poissonconsulting/fwabc/issues).

[Pull requests](https://github.com/poissonconsulting/fwabc/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms
