---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300
)
```
<!-- badges: start -->
[![Lifecycle: deprecated](https://img.shields.io/badge/lifecycle-deprecated-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#deprecated)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

# fwabc

#### ***** This package is no longer being maintained - switched to [fwapgr](https://github.com/poissonconsulting/fwapgr) *****

`fwabc` is an R package to read data from the [BC Freshwater Atlas](https://www2.gov.bc.ca/assets/gov/data/geographic/topography/fwa/fwa_user_guide.pdf). 

`fwabc` uses the [`bcdata`](https://github.com/bcgov/bcdata) package to query the atlas remotely. No local storage of the database is required. `fwabc` provides convenience for common queries and tools/lookups to facilitate working with the atlas.

### Function families 
+ `fwa_read_` - read features from BC Freshwater Atlas layers.
+ `fwa_search_` - match a regular expression to named features.

### Lookup tables
+ `fwa_lookup_gnis` 
+ `fwa_lookup_watershed_group`
+ `fwa_lookup_layer` 

These provide resources for finding layer names, features (GNIS_NAME, WATERSHED_GROUP_NAME, WATERSHED_GROUP_CODE) and which layers are available for given features.

There are 12 variants of `fwa_read_` functions - one for each available layer. These functions are very similar, although arguments and acceptable input differ slightly due to differences in attributes. 

```{r layers, eval = FALSE}
# available layers
fwa_lookup_layer
```

As an example, let's get the stream-network feature for Fraser River. The `fwa_read_stream_network()` accepts a vector of valid GNIS_NAME, WATERSHED_KEY, WATERSHED_GROUP_CODE, or WATERSHED_GROUP_NAME as input. Use the `fwa_search_` functions if unsure about valid input.

```{r stream, eval = FALSE}
fwa_search_gnis("fraser") 

fraser <- fwa_read_stream_network("Fraser River")
fraser


```

Now let's get all named tributaries of the Fraser River with a stream order >= 5.

```{r tribs, eval = FALSE}
fraser_tribs <- fwa_read_stream_network("Fraser River", tributaries = TRUE, named_only = TRUE, min_stream_order = 5L)


```

The equivalent call using the `bcdata` package would be:

```{r bcdata, eval = FALSE}
bcdata::bcdc_query_geodata("92344413-8035-4c08-b996-65a9b3f62fca") %>%
  bcdata::filter(CQL("FWA_WATERSHED_CODE LIKE '100-%'")) %>%
  bcdata::filter(!is.na(GNIS_NAME), STREAM_ORDER >= 5L) %>%
  bcdata::collect()
```

Finally, let's create a map showing a number of layers in the 'Skagit River' watershed group.
```{r map, eval = FALSE}
group <- fwa_search_watershed_group("skagi|scagi")
group

# in this case, 'Skagit River' is both a GNIS_NAME and WATERSHED_GROUP_NAME,
# so we should specify the `input_type`.

stream <- fwa_read_stream_network(group, tributaries = TRUE, 
                                  min_stream_order = 5L, 
                                  input_type = "WATERSHED_GROUP_NAME")
lakes <- fwa_read_lakes(group, input_type = "WATERSHED_GROUP_NAME")
rivers <- fwa_read_rivers(group, input_type = "WATERSHED_GROUP_NAME")
wetlands <- fwa_read_wetlands(group, input_type = "WATERSHED_GROUP_NAME")
border <- fwa_read_watershed_groups(group, input_type = "WATERSHED_GROUP_NAME")

ggplot() +
  geom_sf(data = stream, size = 0.2) + 
  geom_sf(data = lakes, size = 0.3, fill = "steelblue") + 
  geom_sf(data = rivers, size = 0.3, fill = "blue") +
  geom_sf(data = wetlands, size = 0.3, fill = "green")
  geom_sf(data = border, size = 0.07) 


```

<!-- With the `fwabc` package we make these queries easier by doing much of the work in the background and by providing functions and lookup tables to search for valid feature names. For example, the above can be accomplished with: -->

<!-- ```{r bcdata, eval = FALSE} -->
<!-- fwa_search_gnis("fraser river") %>% -->
<!--   fwa_read_stream_network(named_only = TRUE, tributaries = TRUE, min_stream_order = 5L) -->
<!-- ``` -->
<!-- If you prefer to review the result without collecting (which can take a long time for large requests), use the argument `collect = FALSE`. -->

<!-- We currently provide functions for 11 datasets. Since each dataset has different attributes, acceptable x input and function arguments differ slightly for each. For example, fwa_read_watershed_group() accepts only valid WATERSHED_GROUP_CODE or WATERSHED_GROUP_NAME as input and does not provide the option to get tributaries or filter by named/stream order. -->

<!-- If you are familiar with the atlas and require greater query flexibility, then we suggest using the fantastic `bcdata` package directly. -->

## Installation

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
if(!"drat" %in% installed.packages()[,1]) 
  install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("fwabc")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/fwabc)
```
if(!"devtools" %in% installed.packages()[,1]) 
  install.packages("devtools")
devtools::install_github("bcgov/bcdata")
devtools::install_github("poissonconsulting/yesno")
devtools::install_github("poissonconsulting/err")
devtools::install_github("poissonconsulting/checkr")
devtools::install_github("poissonconsulting/fwabc")
```

<!-- ## Usage -->

<!-- ### Function families -->
<!-- + `fwa_read_` - read features from BC Freshwater Atlas layers. -->
<!-- + `fwa_pull_` - return a modified version of some input (e.g. another input type, or tributaries of that input). -->
<!-- + `fwa_search_` - match a regular expression to official named features. -->

<!-- ### Lookup tables -->
<!-- `fwa_lookup_gnis`, `fwa_lookup_watershed_group`, `fwa_lookup_layer` provide resources for finding layer names, named features and which layers have data for particular features.  -->

<!-- ### Read  -->
<!-- Read features from available layers using the `fwa_read()` function: -->
<!-- ```{r stream} -->
<!-- library(ggplot2) -->
<!-- library(magrittr) -->
<!-- library(fwabc) -->

<!-- fwa_read(c("SKGT", 356439092), layer = "stream-network")  -->

<!-- ``` -->
<!-- All `fwa_read_` functions return a [sf](https://github.com/r-spatial/sf) object and require some combination of `WATERSHED_KEY` and `WATERSHED_GROUP_CODE` as input. -->

<!-- ```{r lookup} -->
<!-- fwa_lookup_layer -->
<!-- ``` -->

<!-- Read features from multiple layers for Skagit River watershed group: -->
<!-- ```{r layers} -->
<!-- layers <- c("stream-network", "rivers", "lakes", "watershed-groups") -->
<!-- x <- lapply(layers, function(x) fwa_read("SKGT", layer = x)) -->
<!-- names(x) <- layers -->

<!-- ggplot() +  -->
<!--     geom_sf(data = x[["watershed-groups"]], size = 0.2) + -->
<!--     geom_sf(data = x[["lakes"]], size = 0.3, fill = "steelblue") + -->
<!--     geom_sf(data = x[["rivers"]], size = 0.3, fill = "steelblue") + -->
<!--     geom_sf(data = x[["stream-network"]][x[["stream-network"]]$STREAM_ORDER > 2,], size = 0.07)  -->

<!-- ``` -->

<!-- There is a convenience function for each layer, e.g. `fwa_read_stream_network()`, `fwa_read_watershed_groups()`, `fwa_read_coastlines()` etc. -->
<!-- To read an entire layer, leave the default `x = NULL`. Some layers (e.g. `stream-network` and `watersheds`) are very large. -->

<!-- ```{r wsgroup} -->
<!-- library(rmapshaper) -->
<!-- x <- fwa_read_watershed_groups(ask = FALSE) %>% -->
<!--   # simplify with rmapshaper package -->
<!--   rmapshaper::ms_simplify()   -->

<!-- ggplot() +  -->
<!--   geom_sf(data = x, size = 0.05, aes(fill = WATERSHED_GROUP_NAME), show.legend = FALSE)  -->
<!-- ``` -->

<!-- ### Search and pull -->
<!-- `fwa_search_` and `fwa_pull_` functions can be used to find `WATERSHED_KEY` or `WATERSHED_GROUP_CODE`. -->

<!-- Match a regular expression to an official stream name (`GNIS_NAME`) or watershed group name (`WATERSHED_GROUP_NAME`): -->

<!-- ```{r search} -->
<!-- fwa_search_gnis("skagi|scagi") -->
<!-- fwa_search_watershed_group("skagi|scagi") -->
<!-- ``` -->

<!-- Pull the `WATERSHED_KEY` or `WATERSHED_GROUP_CODE` from the official name and provide to `fwa_read_`: -->

<!-- ```{r pipe} -->
<!-- stream <- fwa_search_gnis("skagi|scagi") %>% -->
<!--   fwa_pull_watershed_key() %>% -->
<!--   fwa_read_stream_network() -->

<!-- wshed <- fwa_search_watershed_group("skagi|scagi") %>% -->
<!--   fwa_pull_watershed_group_code() %>% -->
<!--   fwa_read_watershed_groups() -->

<!-- ggplot() + -->
<!--   geom_sf(data = wshed, size = 0.2) + -->
<!--   geom_sf(data = stream)  -->

<!-- ``` -->

<!-- ### Tributaries -->
<!-- Use `fwa_pull_tributaries()` to get tributaries from a `WATERSHED_KEY`: -->

<!-- ```{r tribs} -->
<!-- tribs <- fwa_search_gnis("skagit") %>% -->
<!--   fwa_pull_watershed_key() %>% -->
<!--   fwa_pull_tributaries(order = 1L) %>% -->
<!--   fwa_read_stream_network() -->

<!-- ggplot() + -->
<!--   geom_sf(data = wshed, size = 0.2) + -->
<!--   geom_sf(data = stream) + -->
<!--   geom_sf(data = tribs, size = 0.1)  -->

<!-- ``` -->

## Contribution

Please report any [issues](https://github.com/poissonconsulting/fwabc/issues).

[Pull requests](https://github.com/poissonconsulting/fwabc/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms
